### 2020-10-22 paper/Automaton.tex
Description du protocol TCP

\begin{figure}[t]
{\centering\scalebox{.7}{
\begin{tikzpicture}[>=latex]
    \def\tfrac#1#2{\ensuremath{\displaystyle\frac{\text{\footnotesize#1}}{\text{\footnotesize#2}}}}
    %
    % Styles for states, and state edges
    %
    \tikzstyle{state} = [draw, very thick, fill=white, rectangle, minimum height=3em, minimum width=7em, node distance=8em, font={\sffamily\bfseries}]
    \tikzstyle{stateEdgePortion} = [black,thick];
    \tikzstyle{stateEdge} = [stateEdgePortion,->];
    \tikzstyle{edgeLabel} = [pos=0.5, text centered, font={\sffamily\small}];

    %
    % Position States
    %
    \node[state, name=closedStart] {CLOSED};
    \node[state, name=listen, below of=closedStart] {LISTEN};
    \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN\_SENT};
    \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RCVD};
    \node[state, name=established, below of=listen, node distance=14em] {ESTABLISHED};
    \node[state, name=finWait1, below of=established, left of=established, node distance=7em, xshift=-9em] {FIN\_WAIT\_1};
    \node[state, name=finWait2, below of=finWait1] {FIN\_WAIT\_2};
    \node[state, name=closeWait, below of=established, right of=established, node distance=7em, xshift=9em] {CLOSE\_WAIT};
    \node[state, name=closing, below of=established, node distance=14em] {CLOSING};
    \node[state, name=lastAck, below of=closeWait] {LAST\_ACK};
    \node[state, name=timeWait, below of=closing] {TIME\_WAIT};

    %
    % Connect States via edges
    %
    \draw ($(closedStart.south) + (-.5em,0)$)
    edge[stateEdge] node[edgeLabel, xshift=-3em]{\tfrac{\emph{Passive open}}{--}}
    ($(listen.north) + (-.5em,0)$);
    \draw ($(listen.north) + (.5em,0)$)
    edge[stateEdge] node[edgeLabel, xshift=2em]{\tfrac{\emph{Close}}{--}}
    ($(closedStart.south) + (.5em,0)$);

    \draw ($(listen.south) + (-1em,0)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=2em]{\tfrac{rcv SYN}{snd SYN, ACK}}
    ($(synRcvd.east) + (0,1em)$);

    % \draw ($(synRcvd.north) + (.5em,0)$)
    % edge[stateEdge, bend left=45] node[edgeLabel,xshift=-4em]{\emph{Timeout}/RST}
    % ($(closedStart.west) + (0,-.5em)$);

    \draw ($(synSent.north) + (-.5em,0)$)
    edge[stateEdge, bend right=45] node[edgeLabel,xshift=-1em, yshift=-2em]{\tfrac{\emph{Close}}{--}}
    ($(closedStart.east) + (0,-.5em)$);
    \draw ($(closedStart.east) + (0,.5em)$)
    edge[stateEdge, bend left=45] node[edgeLabel,xshift=4em]{\tfrac{\emph{Active open}}{snd SYN}}
    ($(synSent.north) + (.5em,0)$);

    \draw (synSent.west)
    edge[stateEdge] node[edgeLabel, yshift=1.3em]{\tfrac{rcv SYN}{snd SYN, ACK}}
    (synRcvd.east);
    \draw (synRcvd)
    edge[stateEdge] node[edgeLabel, xshift=-2.5em]{\tfrac{\emph{Close}}{snd FIN}}
    (finWait1);

    \draw ($(synRcvd.east) + (0,-1em)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1.5em, yshift=-2em]{\tfrac{rcv ACK}{--}}
    ($(established.north) + (-1em,0)$);
    \draw ($(synSent.west) + (0,-1em)$)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{\tfrac{rcv SYN, ACK}{snd ACK}}
    ($(established.north) + (1em,0)$);

    \draw ($(established.south) + (-1em,0)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=1.5em]{\tfrac{\emph{Close}}{snd FIN}}
    ($(finWait1.east) + (0,.5em)$);
    \draw ($(established.south) + (1em,0)$)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1.5em]{\tfrac{rcv FIN}{snd ACK}}
    ($(closeWait.west) + (0,1em)$);

    \draw (finWait1.south)
    edge[stateEdge] node[edgeLabel, xshift=-2em]{\tfrac{rcv ACK}{--}}
    (finWait2.north);
    \draw ($(finWait1.east) + (0,-.5em)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=0em, xshift=4.5em]{\tfrac{rcv FIN}{snd ACK}}
    (closing.north);
    \draw (finWait1.south east)
    edge[stateEdge] node[edgeLabel, xshift=0em, yshift=2em, text width=3em]{\tfrac{rcv FIN, ACK}{snd ACK}}
    (timeWait.north west);

    \draw (finWait2.south)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=-2em, yshift=-1em]{\tfrac{rcv FIN}{snd ACK}}
    (timeWait.west);

    \draw (closing)
    edge[stateEdge] node[edgeLabel, xshift=2.5em]{\tfrac{rcv ACK}{--}}
    (timeWait);

    \draw (closeWait)
    edge[stateEdge] node[edgeLabel,xshift=-2.5em]{\tfrac{\emph{Close}}{snd FIN}}
    (lastAck);

    %
    % Connect lastAck to closed is slightly more complicated
    % no direct line-of-sight, so we need to take the scenic route
    %
    \coordinate (lastAck2ClosedA) at ($(lastAck.east) + (2em,0)$);
    \coordinate (lastAck2ClosedB) at ($(closedStart.north -| lastAck.east) + (2em,1em)$);
    \coordinate (lastAck2ClosedC) at ($(closedStart.north) + (0.5em,1em)$);
    \draw (lastAck.east) edge[stateEdgePortion] (lastAck2ClosedA);
    \draw (lastAck2ClosedA) edge[stateEdgePortion] node[edgeLabel,xshift=-2.5em, yshift=-4em]{\tfrac{rcv ACK}{--}} (lastAck2ClosedB);
    \draw (lastAck2ClosedB) edge[stateEdgePortion] (lastAck2ClosedC);
    \draw (lastAck2ClosedC) edge[stateEdge] ($(closedStart.north) + (0.5em,0)$);

    %
    % likewise for timeWait to closed
    %
    \coordinate (timeWait2ClosedA) at ($(timeWait.south) + (0,-1em)$);
    \coordinate (timeWait2ClosedB) at ($(timeWait.south -| finWait2.west) + (-2em,-1em)$);
    \coordinate (timeWait2ClosedC) at ($(closedStart.north -| finWait2.west) + (-2em,1em)$);
    \coordinate (timeWait2ClosedD) at ($(closedStart.north) + (-0.5em,1em)$);
    \draw (timeWait.south) edge[stateEdgePortion] (timeWait2ClosedA);
    \draw (timeWait2ClosedA) edge[stateEdgePortion] (timeWait2ClosedB);
    \draw (timeWait2ClosedB) edge[stateEdgePortion] (timeWait2ClosedC);
    \draw (timeWait2ClosedC) edge[stateEdgePortion]
    node[edgeLabel, text width=12.25em, yshift=1.5em]{\emph{Timeout after two maximum segment lifetimes (2*MSL)}}
    (timeWait2ClosedD);
    \draw (timeWait2ClosedD) edge[stateEdge] ($(closedStart.north) + (-0.5em,0)$);

    % draw dotted lines around passive and active closes
    % \begin{pgfonlayer}{background}
    %     \draw [join=round,black,dotted] ($(closeWait.north west) + (-1em, -1em)$) rectangle ($(lastAck.south east) + (1em, 1em)$);
    %     \draw [join=round,black,dotted] ($(finWait1.north west) + (-1em, -1em)$) rectangle ($(timeWait.south east) + (1em, 1em)$);
    % \end{pgfonlayer}

\end{tikzpicture}}\par}
\caption{TCP state machine.}
\label{Fig:statemachine}
\end{figure}