### 2021-05-24 SecDev21/paper.tex
rephrasing

\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage[x11names]{xcolor}

\usepackage[np]{numprint}

\definecolor{linkcolor}{RGB}{36,113,163}
\definecolor{citecolor}{RGB}{17,122,101}
\definecolor{urlcolor}{RGB}{148,39,38}

\usepackage[
    colorlinks,
    linkcolor=linkcolor,
    citecolor=citecolor,
    urlcolor=urlcolor
]{hyperref}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{amsmath, amsfonts}

\usepackage{lipsum}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing, patterns}
\usetikzlibrary{calc, backgrounds}

\usepackage{xspace}

\usepackage{listings}
\usepackage{cleveref}

\lstset{
    basicstyle=\ttfamily,
    keywordstyle={[1]\color{Blue4}\bfseries}, % keywords style
    % keywordstyle={[2]\color{Pink4}}, % type style
    % keywordstyle={[3]\color{OrangeRed3}}, % function style
    % literate =
    %     {=>}{{=>}}2
    %     {/=}{{/=}}2,
    commentstyle=\itshape\color{gray}
}

\def\spark#1{\lstinline[language=Ada]{#1}}

\usepackage{todonotes}
%\def\todo#1{\textcolor{red}{[#1]}}

\def\state#1{\textsf{\MakeUppercase{#1}}\xspace}
\def\sclosed{\state{closed}}
\def\ssynsent{\state{syn-sent}}
\def\ssynrcv{\state{syn-rcvd}}
\def\slisten{\state{listen}}
\def\sestab{\state{established}}
\def\sfwone{\state{fin-wait-1}}
\def\sfwtwo{\state{fin-wait-2}}
\def\sclosing{\state{closing}}
\def\sclosew{\state{close-wait}}
\def\slastack{\state{last-ack}}
\def\stimewait{\state{time-wait}}

\def\flag#1{\textsf{#1}\xspace}
\def\syn{\flag{SYN}}
\def\ack{\flag{ACK}}
\def\rst{\flag{RST}}
\def\fin{\flag{FIN}}

\let\code\texttt

\begin{document}

\title{Layered Formal Verification of an Industrial TCP/IP Stack}

\author{%
\IEEEauthorblockN{Guillaume Cluzel}
\IEEEauthorblockA{\textit{AdaCore \& ENS de Lyon}}
\and
\IEEEauthorblockN{Kyriakos Georgious}
\IEEEauthorblockA{\textit{AdaCore \& University of Bristol}}
\and
\IEEEauthorblockN{Yannick Moy}
\IEEEauthorblockA{\textit{AdaCore}}
\and
\IEEEauthorblockN{Clément Zeller}
\IEEEauthorblockA{\textit{Oryx Embedded}}
}

\maketitle

\begin{abstract}
\todo{Write the abstract.}
\end{abstract}

\begin{IEEEkeywords}
Network protocols,
deductive verification.
\end{IEEEkeywords}

\section{Introduction}

By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}. At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}. Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices. Researchers have shown that such vulnerabilities exist: in 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.

Networking communication models define the processes of transferring information from one network component to another. Currently, the two most popular networking communication models are the TCP/IP (Transmission Control Protocol/Internet Protocol) \cite{TCP_IP:ietf_tutorial} and the OSI (Open System Interconnection) \cite{ISO:35.100} models. Layering is an essential design requirement for both models to achieve modularity, flexibility, and abstraction. Applications that only need to use the lower levels' functionality can drop all the unnecessary upper levels. \Cref{Fig:TcpStack} shows the various layers of the OSI model. Each layer supports a wide range of protocols that are compatible with the layer's specifications. The lowest layers, bottom layers in \Cref{Fig:TcpStack}, are more hardware-oriented, while the top layers are more software-specific. While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP). Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Application layers. %Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.

\input{OryxStack.tex}

Existing research has dealt with formally verifying several protocols of TCP/IP stack levels. Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}. More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL~\cite{tcp2005sigcomm,ridge2008rigorous} and used the resulting specification for testing existing libraries. Others have formally specified and implemented the SSL/TLS protocol in F*\cite{bhargavan2013implementing}, and integrated the resulting implementation in existing libraries. To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and Socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications~\cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.

\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}

We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

\section{Overview of the TCP protocol specification}
\label{sec:TCP}

\input{Automaton.tex}

TCP is a connection-oriented protocol; a connection between the sender and the receiver must be established before transmitting any data. Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order they were sent. It also provides error-checking mechanisms to discard and recover from corrupted data.

%\subsection{TCP state machine}

Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}. An edge represents a state transition.
An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated to the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response to the action that caused the state transition, when present.
An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event. The flags are embedded in the header section of a transmitted segment and can be one of the following:
\begin{itemize}
\item \ack: the last message received by the sender is acknowledged.
\item \syn: this flag is sent to establish a connection.
\item \fin: this flag is sent to close the connection.
\item \rst: this flag is sent to reset the connection, often when an error
occurred on one or the other side.
\end{itemize}

The state machine in \cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes. Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime. Each state has a meaning. \slisten represents waiting for a connection request from any remote TCP and port. \ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached. \sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represent waiting for a connection termination request from the remote TCP. \sclosew and \slastack represent waiting for a connection termination request from the local user. Finally, \sclosed represents no connection state at all.

%\subsection{TCP Multitasking model}

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. RFC 793, under Section ``3.9. Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follows. User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, corresponding transitions to the timeout events can also be triggered by this task. A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.

The CycloneTCP implementation adopts the three-task-based model and the socket data structure defined by the RFC 793. Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of incoming segments. Moreover, the user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.

\section{Specification techniques used}
\label{sec:spec}

\subsection{The SPARK programming language}

Ada is a general-purpose, procedural programming language that puts great emphasis on the safety and correctness of the program. Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks. Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings cannot exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers. Furthermore, Ada introduced contracts, such as preconditions and postconditions, as part of the language’s standard syntax. Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.

By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} emerged to provide the largest possible subset of Ada, which is suitable for functional specification and static verification. For instance, in SPARK, expressions should be free from side effects, and aliasing is forbidden (no two variables should share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.

\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses. The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detects unused assignments and unmodified variables. The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.

\subsection{Interfacing SPARK and C code}

The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work. Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API). This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.

Ada and SPARK offer an easy-to-use mechanism for interfacing with C code \cite{}. The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects. The memory representations of a shared data object must be the same in both the C and the SPARK code. For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument. It is the responsibility of the user to ensure this.

Although formal verification can only be applied to the SPARK translated code,  preconditions and postconditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code. Since the functional specification of these contracts cannot be proved, it is the programmer's role to ensure their correctness. Interfacing C and SPARK code can be error-prone if it is not done carefully. \Cref{sec:verif} discusses how to mitigate this issue effectively.

\todo[inline]{KG: fixed up to here}

\subsection{Dealing with pointers}
\label{sec:pointers}

In the CycloneTCP library, pointers are used both as the type of variables, to pass objects when calling other functions, and the type of structure fields, for instance in
the socket structure to store other complex data-structures and recursive pointer-based data-structures. SPARK use a pointer ownership model to prevent bugs
involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are manipulated.
When a memory area is allocated, a pointer points to this memory area and thus it is owned.
The ownership of the memory can be moved but the memory must be deallocated
before the end of the execution of the program.

\subsection{Specifying the frame condition}

The socket structure is shared between different functions, and contains lots
of fields. Most of the fields are used in other parts of the library and are not
relevant for our verification. Thus we use \emph{ghost code} in SPARK
to focus the verification effort on the relevant subset of the fields,
which also makes the specifications easier to write and to understand.
This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects.
Ghost code is used by GNATprove for analysis but not compiled in the final executable.

More specifically, we define a \emph{model} of the actual socket as a ghost function for use in contracts: \spark{Model(Sock)}
extracts the fields of interest from socket \spark{Sock}.
A frequent usage in postconditions is
\begin{lstlisting}[language=Ada, basicstyle=\small\ttfamily]
Model(Sock) = (Model(Sock)'Old with delta
                    State => TCP_STATE_CLOSED)
\end{lstlisting}
that states that relevant fields of \spark{Sock} selected by \spark{Model} are
not changed by the function call, except for the field \spark{State} that is equal
to \spark{TCP_STATE_CLOSED} after the call.

\section{Conformance to the TCP protocol}
\label{sec:verif}

% The reader can refer to the code at \url{https://github.com/AdaCore/Http_Cyclone}.

\subsection{Extracting a specification from RFC 793}

Extracting a specification
is mandatory for provide functional contracts to the user functions. That can
be achieved by reading the text of RFC 793. However, the text is underspecified in many
cases and only one possible implementation is described. We made the following choices in our formalization of the TCP specification:
\begin{itemize}
\item The transitions between states must respect the order given by the
state machine described in figure~\ref{Fig:statemachine}.
In particular, a transition from a state to
another is valid only if there exists a transition in the state machine between
these two states.
\item RFC 793 describes conditions on the state of the socket for calling user functions,
which need to hold when the function is called~\cite[p. 52]{rfc793}. In particular it is specified in what state the
function can be called without returning an error. These became preconditions in our formalization.
\end{itemize}

\subsection{Rewriting the TCP user functions}

The user functions defining the API of the TCP protocol were fully rewritten in SPARK.
SPARK bindings were also provided to the majority of the rest of the library’s C functions to allow their invocation from the SPARK code,
taking into account their effect on the TCP state machine in their contracts.
GNATprove was used to prove that the implementation of user functions is free of run-time errors and
respects the functional specifications of the TCP protocol that we formalized in contracts.

% dire que les fonctions appelées dans le code on était modélisées pour prendre
% en compte les changements d'état TCP qu'elles occasionnent.

% Problème, certains changement d'état se font entre l'appel de deux fonctions,
% ça doit donc être pris en compte. Et certains état se font directement lors de
% l'appel d'une fonction particulière et ça doit aussi être pris en compte.

To model transitions of the TCP state machine inside the code, we defined the helper
function \spark{TCP_Change_State}, whose contract describes
all the transitions allowed.
As calls to user functions are protected with a mutex to avoid interferences
from other threads, the only changes to the TCP state machine inside user functions are
related to their actions. We model these changes by calling \spark{TCP_Change_State}
at places that correspond to a transition of the TCP machine.

% Yannick: I don't understand the following, needs to be rephrased if we want to keep it
%
% particular when a new segment is received, it is quite easy to know the
% possible set of state that can have the socket during the all execution of the
% functions.

The function \spark{TCP_Wait_For_Events} requires special handling because it may
release the mutex. This function takes in argument an event on which it should wait,
and returns once the event is triggered by the reception of a segment.
While waiting, it releases the mutex to allow the reception task to process incoming segments.
For example, this function is called from \spark{open} after the connection
has been initialized, to wait until the connection has been established or not, in order to
correctly set the error that indicates to the user the state of the connection.

% Yannick: we talk about error above, but there is no error if the connection is established???
% This needs rephrasing

A correct but imprecise contract for \spark{TCP_Wait_For_Events} would allow any change in the TCP
state machine while waiting. But it would make it impossible to prove that its callers
satisfy their functional specification. We need a more precise contract for this function.

\subsection{Dealing with synchronous changes of states}

% Yannick: below "a new segement...if it contains a new segment" is not phrased correctly

The function \spark{TCP_Wait_For_Events} first checks if the event is not already
true. If not, the mutex is released until a segment that triggered the event is
received or until the timeout is reached. Each time a new segment is received, it
is processed and if it contains a segment that causes a change of state, the TCP
state is updated. Then the events are updated, and if the waited event becomes
true due to the change of state, it is triggered.

% Yannick: the above final sentence is unclear: what is triggered?

\begin{algorithm}[t]
    \caption{Function to compute the possible state after the completion of a
particular event that is requested by a user-task related function.}
\label{algo:waitForEvents}
\begin{algorithmic}[1]
\footnotesize
\Function{TCP\_Wait\_For\_Events\_Proof}{Socket, Event\_Mask}
    \State $S_{last} := \text{Socket}$
    \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
    \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
        \State \textbf{return}  $S_{last}$
    \EndIf
    \For{$i=1$ \textbf{to} $3$}
        \State $S_{last} :=$ \spark{TCP_Process_One_Segment}($S_{last}$)
        \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
        \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
            \State \textbf{return} $S_{last}$
        \EndIf
    \EndFor
    \State \textbf{return} $\emptyset$
\EndFunction
\end{algorithmic}
\end{algorithm}

An event is considered to be true or not depending on the state of the socket.
This computation is done by the function \spark{TCP_Update_Events}. We have put
a contract on it, which is proved to be correct by SPARK. Then we can write a
function that emulates what is done when \spark{TCP_Wait_For_Events} is called.
We have written a procedure \spark{TCP_Wait_For_Events_Proof}, reproduced in
algorithm~\ref{algo:waitForEvents} to compute all the possible TCP states that a
socket can have after a wait. This algorithm simulates the reception of
one message, update the events and return if the event waited is reached.
Otherwise it repeats the procedures. The for loop only requires three steps
because all the step in the state machine can be reached in three segment
receptions without user interaction.

On \spark{TCP_Wait_For_Events_Proof}, a contract has been put to describe all
the possible states that a socket can have after the wait. It is only possible
to prove this contract with SPARK if the function named
\spark{TCP_Process_One_Segment} in algorithm~\ref{algo:waitForEvents}
has a postcondition.

\subsection{Symbolic execution to extract contracts}

Rewriting the part that process the incoming segments was out of the scope of
this work. Instead, we have opted for the use of symbolic execution on the
original CycloneTCP C code.
Symbolic execution is a mean to execute programs with symbolic values rather
than concrete values. Symbolic execution has already been successfully
used with other verification methods.
Indeed, in~\cite{vanoverberghe2008using} the authors present how to use symbolic
execution to improve deductive verification, and in~\cite{kassios2012comparing}
the authors explain that symbolic execution can be good at infering contracts.
For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM
bytecode in input and works well with C code thanks to a very simple interface.

\begin{figure}
\begin{lstlisting}[language=C, basicstyle=\footnotesize\ttfamily, numbers=left,
    numberstyle=\tiny, frame=bottomline, escapeinside={(*@}{@*)}]
// creation of a fake incoming segment
TCPheader *segment = malloc(sizeof(TCPheader));
klee_make_symbolic(segment, sizeof(segment), "seg");
klee_assume(segment->flag <= 31);
// Creation of the socket
Socket *sock = malloc(sizeof(Socket)), oldSock;
klee_make_symbolic(sock, sizeof(sock), "sock");
memcpy(&oldSock, sock, sizeof(Socket));

// The function that process the segments is called
tcpProcessSegment(sock, segment); (*@\label{code:kleedriver:tcpProcessSegment}@*)

// Contract to check
klee_assert( (*@\label{code:kleedriver:assert}@*)
    (oldSock.state == TCP_STATE_ESTABLISHED) ?
        sock.state == TCP_STATE_ESTABLISHED ||
        sock.state == TCP_STATE_CLOSE_WAIT ||
        sock.state == TCP_STATE_CLOSED :
    (oldSock.state == ...) ? ... )
)
\end{lstlisting}
\caption{Driver for the verification of \lstinline[language=C]{tcpProcessSegment}
with KLEE}
\label{code:kleedriver}
\end{figure}

The first step in order to extract contracts was to read the state machine to
know all the possible transitions that can be done by the reception of a flag,
and adapt it to
the subtleties of CycloneTCP. The next step is to write a driver to run KLEE and
know if the contracts that we have infered are correct or not.
A simplified example is shown in figure~\ref{code:kleedriver}. The first lines
create a symbolic incoming segment to process as well as a symbolic socket.
At line~\ref{code:kleedriver:tcpProcessSegment} the function that processes the
segment is called. It is followed at line~\ref{code:kleedriver:assert} by a KLEE
builtin that check if an assert is valid for all the symbolic paths
explored before.

Once the contract is validated by KLEE, it can be copied out on the SPARK
function \spark{TCP_Process_One_Segment}. With this step done, the concurrency
challenge has been solved and the function
\spark{TCP_Wait_For_Events} have a correct and an accurate contract, which
makes possible the verification of the functional correctness of the TCP user
functions thanks to SPARK.

\section{Hardening the user's API}
\label{sec:API}

It appears that lot of programmers do not correctly use the socket API that is
provided by the library. Providing verified user functions for TCP is not enough
to guarantee the safety of software if the users do not correctly use them.
An important part of the work was to ensure that code that uses the socket API's
functions are correctly called. By ``correctly'' we mean: 1)~the functions
are called with respect to a defined order and 2)~the return code is checked
before continuing the processing.

\subsection{Enforcing the correct order of API functions}

The TCP protocol implies a specific order such that the user can call the API's
functions without breaking the functional specification of the protocol. This
order also is being conveyed by the protocol's state machine given in
figure~\ref{Fig:statemachine}.
Ada's pre- and postconditions are a powerful tool to express such inter-function
dependencies, while SPARK technology can be used to guarantee that the
assertions always hold. Thus, pre- and postconditions were introduced to model
a partial order on the calls to the API's functions.

\begin{figure}
\begin{lstlisting}[language=Ada,basicstyle=\footnotesize\ttfamily]
procedure Socket_Connect
   (Sock           : in out Not_Null_Socket;
    Remote_Ip_Addr : in     IpAddr;
    Remote_Port    : in     Port;
    Error          :    out Error_T)
   with
   Pre => Is_Initialized_Ip (Remote_Ip_Addr),
   Post =>
    if Sock.S_Type = SOCKET_TYPE_STREAM then
      (if Error = NO_ERROR then
        Sock.S_Remote_Ip_Addr = Remote_Ip_Addr)
    else
      Sock.S_Remote_Ip_Addr = Remote_Ip_Addr

procedure Socket_Send
   (Sock    : in out Not_Null_Socket;
    Data    : in     Send_Buffer;
    Written :    out Natural;
    Flags   :        Socket_Flags;
    Error   :    out Error_T)
   with
   Pre => Is_Initialized_Ip (Sock.S_Remote_Ip_Addr)
\end{lstlisting}
\caption{An example of how function calls can be ordered by pre- and
postconditions}
\label{fig:functionorder}
\end{figure}

Simplified contracts extracted from our library are reproduced in
figure~\ref{fig:functionorder}. If no error occurs, the function
\spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to
\spark{Remote_Ip_Addr} which is supposed to be
initialized when the function is called. \spark{Socket_Send} requires to be
called with a socket not null such that its field \spark{Remote_Ip_Addr} is
initialized. The only way to ensure that the precondition hold is to
call \spark{Socket_Connect} before \spark{Socket_Send}.
The tool GNATprove can statically find if a call to \spark{Socket_Send}
is not preceded by a call to \spark{Socket_Connect}.

\subsection{Checking the correctness of return codes}

A common mistake is to forget to check the return codes after a call to a socket
API's function. For instance, the following code would be reject by SPARK:
\begin{lstlisting}[language=Ada,basicstyle=\small\ttfamily]
Socket_Connect(Sock, Ip_Addr, Port, Error);
Socket_Send(Sock, Data, Written, Flags, Error);
\end{lstlisting}
Indeed, if \spark{Socket_Connect} fails and \spark{Error} is not equal to
\spark{NO_ERROR}, \spark{Sock.S_Remote_Ip_Addr} might be non initialized, and
then the function \spark{Socket_Send} cannot be called because of its
precondition.

\section{Results}
\label{sec:results}

\subsection{Bugs found}

\subsubsection{Memory leak}

Thanks to the pointer ownership policy adopted by SPARK and detailed in
section~\ref{sec:pointers}, we have found that a memory leak can happen when
the connection have been closed and the buffer are cleaning.

\subsubsection{Violation of the TCP protocol}

Another very interesting and very subtle bug have been found thanks to SPARK is
reproduced in the snippet bellow.

\begin{lstlisting}[language=Ada, basicstyle=\footnotesize\ttfamily,
                    numbers=left, numberstyle=\tiny, escapechar=\%]
case Sock.State is
  when TCP_STATE_SYN_RECEIVED
     | TCP_STATE_ESTABLISHED =>
  -- Flush the send buffer
  TCP_Send (Sock, Buf, Ignore_Written,
            SOCKET_FLAG_NO_DELAY, Error);
  if Error /= NO_ERROR then
      return;
  end if;

  -- Make sure all the data has been sent out
  TCP_Wait_For_Events % \label{bugProg:tcpWaitForEvents} %
     (Sock       => Sock,
      Event_Mask => SOCKET_EVENT_TX_DONE,
      Timeout    => Sock.S_Timeout,
      Event      => Event);

  -- Timeout error?
  if Event /= SOCKET_EVENT_TX_DONE then
     Error := ERROR_TIMEOUT;
     return;
  end if;

  -- Send a FIN segment
  TCP_Send_Segment   % \label{bugProg:tcpWaitForEventsPost} %
     (Sock         => Sock,
      Flags        => TCP_FLAG_FIN or TCP_FLAG_ACK,
      Seq_Num      => Sock.sndNxt,
      Ack_Num      => Sock.rcvNxt,
      Length       => 0,
      Add_To_Queue => True,
      Error        => Error);

  -- Failed to send FIN segment?
  if Error /= NO_ERROR then
     return;
  end if;

  -- Switch to the FIN-WAIT-1 state
  TCP_Change_State (Sock, TCP_STATE_FIN_WAIT_1); % \label{bugProg:tcpChangeState} %
\end{lstlisting}

In this scenario, after the call to \spark{Tcp_Send}, when the error cases have
been filter, the state of the socket can either be \sestab or \sclosew. The call
at line~\ref{bugProg:tcpWaitForEvents} and now the state of the socket can be
still be \sestab or \sclosew, or if a \rst flag has been received when the mutex
was released, the state can be \sclosed too.
At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does
not modify the state of the socket. Finally, at line~\ref{bugProg:tcpChangeState}
we try to jump to the state \sfwone and thus the transitions $\text{\sclosew}
\rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are
possible.

But these transition are not allowed by the norm, and thus not allowed by the
precondition of \spark{TCP_Change_State}. The problem is detected by SPARK and a fix
has been proposed.

\subsection{Verified parts}

For achieving this verification, 50 subprograms have been translated from C to
SPARK. This accounts for \np{2266} lines of code of which \np{1165} are logical
code to write and prove the specifications.

\section{Conclusion and future work}

The use of SPARK for the verification of the TCP protocol has given good
results. It has helped to find bugs in the existing implementation. But more
than finding bugs, the interest of this work lie in the face that it is a big
step toward a secure implementation. It has been proved by the tool GNATprove
that our implementation is free of run-time errors and that all the transitions
in the TCP state machine are done with respect to the norm.

However weeknesses exist in our implementation, in particular because all the
underlying layers are still written in C. The principal function of those layers
is to format packet before they are sent, or parse incoming packets, check their
integrity and transmit the resulting payload to the corresponding upper level
layer. This processing part can be a source of errors and bugs as explain
in~\cite{Reiher2019RecordFluxFM}.
Using the RecordFlux DSL to parse the packets of the different protocols is the
next step to make the stack safer. Using RecordFlux to parse TCP message
would be a big step to finish the translation of the TCP protocol in SPARK.
The postconditions extracted by KLEE could be proved by GNATprove for a
safer result.

Our verification of the TCP protocol only focuses on the validity of the
transitions in the TCP state machine, but some other properties are formulated
in RFC 793, and could be verified in the implementation to make it even
more robust.

\bibliographystyle{IEEEtran}
\bibliography{biblio}

\end{document}